---
title: "Genetic Gain"
format: html
editor: visual
---

```{r, include = F}
library(dplyr)
```

### 1. Heritability

```{r}
reps <- 3
years <- 4
locations <- 1
env <- years * locations


anova_1 <- 
data.frame(SV = c("Y", "R/Y", "F", "FxY", "R/YxF", "Plants/plot"), 
           df = c(3,8,99,297,792,10800), 
           MS = c(3190,47,2685,527,92,31)
           ) %>% 
  mutate(VAR = MS * df)
anova_1
```

Heritability or $h^2$ is the ratio between the genotipic variance $\sigma^2_g$ and the phenotypic variance $\sigma^2_{ph}$

$$h^2 = \frac{\sigma^2_g}{\sigma^2_{ph}}$$



Depending on the method $\sigma^2_{ph}$ is estimated differently:

#### 1a. On a single plant basis
  
$$\sigma_{ph}^2 = \sigma^2_{w} + \sigma^2_{ge} + \sigma^2_{g}$$  
$\sigma^2_{w}$ is the variance among plants within a plot and $\sigma^2_{ge}$ is the variance of the genotype by environment interaction.   


```{r}
s2_w <- anova_1$VAR[[6]]
s2_ge <- anova_1$VAR[[4]]
s2_g <- anova_1$VAR[[3]]
h2_a <- s2_g / (sum(s2_w, s2_ge, s2_g ))
h2_a
```


#### 1b. Selection in based on a comparison among plots
  
$$\sigma_{ph}^2 = \sigma^2_{w} + \sigma^2_{plants} + \sigma^2_{ge} + \sigma^2_{g}$$
where $\sigma^2_{plants} = \frac{\sigma^2_{R/Y~x~G}-\sigma^2_w}{N}$ and $N$ is the number of plants per plot. 

```{r}
s2_RYxF <- anova_1$VAR[[5]]

# N
plants_plot <- 10

s2_plants <- (s2_RYxF - s2_w) / plants_plot

h2_b <-  s2_g / (sum(s2_w,s2_plants, s2_ge, s2_g ))
h2_b
```

  
#### 1c. Plot basis

$$\sigma_{ph}^2 = \frac{\sigma^2_{w}}{N} + \sigma^2_{plants} + \sigma^2_{ge} + \sigma^2_{g}$$
  
where $\sigma^2_e = \frac{\sigma^2_w}{N}+\sigma^2_{plants}$ and $N$ is the number of plants per plot. 

```{r}
h2_c <-  s2_g / (sum(s2_w/plants_plot,s2_plants, s2_ge, s2_g ))
h2_c
```


#### 1d. Entry-mean basis
  
$$\sigma^2_{ph} = \frac{\sigma^2_e}{Reps~ Env} + \frac{\sigma^2_{ge}}{Env} + \sigma^2_g$$


```{r}
s2_e <- s2_w/plants_plot + s2_plants
h2_d <-  s2_g / (sum(s2_e/(reps * env), s2_ge/env, s2_g ))
h2_d
```


