---
title: "Genetic Gain"
format: html
editor: visual
---

```{r, include = F}
library(dplyr)
```

### 1. Heritability

```{r}
reps <- 3
years <- 4
locations <- 1
env <- years * locations
N <- 10

anova_1 <- 
data.frame(SV = c("Y", "R/Y", "F", "FxY", "R/YxF", "Plants/plot"), 
           df = c(3,8,99,297,792,10800), 
           MS = c(3190,47,2685,527,92,31)
           )
anova_1
```

```{r}
# Get variance components from ANOVA table
s2_w <- anova_1$MS[[6]]
s2 <- (anova_1$MS[[5]] - s2_w)/N
s2_ge <- (anova_1$MS[[4]] - (s2_w + N*s2))/(N*reps)
s2_g <- (anova_1$MS[[3]] - (s2_w + N*s2 + N*reps*s2_ge))/(N*reps*env)
```


Heritability or $h^2$ is the ratio between the genotipic variance $\sigma^2_g$ and the phenotypic variance $\sigma^2_{ph}$

$$h^2 = \frac{\sigma^2_g}{\sigma^2_{ph}}$$



Depending on the method $\sigma^2_{ph}$ is estimated differently:

#### 1a. On a single plant basis
  
$$\sigma_{ph}^2 = \sigma^2_{w} + \sigma^2_{ge} + \sigma^2_{g}$$  
$\sigma^2_{w}$ is the variance among plants within a plot and $\sigma^2_{ge}$ is the variance of the genotype by environment interaction.   


```{r}
h2_a <- s2_g / (sum(s2_w,s2, s2_ge, s2_g))
h2_a
```


#### 1b. Selection in based on a comparison among plots
  
$$\sigma_{ph}^2 = \sigma^2_{w} + \sigma^2 + \sigma^2_{ge} + \sigma^2_{g}$$

```{r}
h2_b <-  s2_g / (sum(s2_w, s2_ge, s2_g ))
h2_b
```

  
#### 1c. Plot basis

$$\sigma_{ph}^2 = \frac{\sigma^2_{w}}{N} + \sigma^2 + \sigma^2_{ge} + \sigma^2_{g}$$
  

```{r}
h2_c <-  s2_g / (sum(s2_w/N,s2, s2_ge, s2_g ))
h2_c
```


#### 1d. Entry-mean basis
  
$$\sigma^2_{ph} = \frac{\sigma^2_e}{Reps~ Env} + \frac{\sigma^2_{ge}}{Env} + \sigma^2_g$$
where $\sigma^2_e = \frac{\sigma^2_w}{N}+\sigma^2$ and $N$ is the number of plants per plot. 


```{r}
s2_e <- s2_w/N + s2
h2_d <-  s2_g / (sum(s2_e/(reps * env), s2_ge/env, s2_g ))
h2_d
```

### 2. Genetic Gain

$$G_{cycle} = \frac{c~k~\sigma^2_g}{\sigma^2_{ph}}$$

$$G_{year} = \frac{G_{cycle}}{Y} $$
where $Y$ is the number of years that takes to complete a cycle of selection.

For this exercise $k  = 1.40$


#### 2.1 

Recurrent phenotypic selection. Female control only, selection within plots (grid selection).

$$G_{cycle} = \frac{k~c~\sigma^2_A}{ \sqrt{\sigma^2_{ph}}}$$
where $\sigma^2_{ph} = \sigma^2_{w} + \sigma^2_{ge} + \sigma^2_{g}$

```{r}
c <- 0.5
k <- 1.40
s2_ph <- sum(s2_w, s2_ge, s2_g)
gc_1 <- k * c * s2_g / sqrt(s2_ph)
gc_1
```
#### 2.2

Modified ear‑to‑row. No selection within rows.

$$G_{cycle} = \frac{k~c~\frac{1}{4}\sigma^2_A}{\sqrt{\sigma^2_{ph}}}$$

$$\sigma^2_{ph} = \frac{\sigma^2_{e}}{Reps ~ Env} + \frac{\frac{1}{4} \sigma^2_{ge}}{Env} + \sigma^2_{g}$$

```{r}
c <- 0.5
k <- 1.40
s2_ph <- sum(s2_e/(reps*env), (0.25* s2_ge)/env, s2_g)
gc_1 <- k * c * 0.25 * s2_g / sqrt(s2_ph)
gc_1
```

#### 2.3

Half‑sib selection, population as tester, recombine remnant half‑sib seed.  Male and female control.

$$G_{cycle} = \frac{k~c~\frac{1}{4}\sigma^2_A}{\sqrt{\sigma^2_{ph}}}$$

$$\sigma^2_{ph} = \frac{\sigma^2_{e}}{Reps ~ Env} + \frac{\frac{1}{4} \sigma^2_{ge}}{Env} + \sigma^2_{g}$$

```{r}
# Male and female control
c <- 1
k <- 1.40
s2_ph <- sum(s2_e/(reps*env), (0.25* s2_ge)/env, s2_g)
gc_1 <- k * c * 0.25 * s2_g / sqrt(s2_ph)
gc_1
```

#### 2.4

Half‑sib selection, population tester, recombine selfed seed.

$$G_{cycle} = \frac{k~c~\frac{1}{4}\sigma^2_A}{\sqrt{\sigma^2_{ph}}}$$

$$\sigma^2_{ph} = \frac{\sigma^2_{e}}{Reps ~ Env} + \frac{\frac{1}{4} \sigma^2_{ge}}{Env} + \sigma^2_{g}$$

```{r}
# Male and female control
c <- 2
k <- 1.40
s2_ph <- sum(s2_e/(reps*env), (0.25* s2_ge)/env, s2_g)
gc_1 <- k * c * 0.25 * s2_g / sqrt(s2_ph)
gc_1
```



